### 2021-05-21 papers/SYRCoSE/future.tex
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

* feat: add part of small-step

First, we plan to apply our library to a wider range of problems. 
We are going to develop a mechanised semantics of some long-established
languages used to model concurrency, in particular 
the calculus of communicating systems~(CCS)~\cite{Milner:80} and 
$\pi$-calculus~\cite{Milner:99}.      
We also plan to continue our work on 
expressing various relaxed models 
of shared memory~\cite{Lahav-al:POPL16, Lahav-al:PLDI17, Podkopaev-al:POPL19}
in terms of event structures.

Second, we want to cover other classes of event structures in our library, 
in particular bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991}
and stable~\cite{Winskel:82, Winskel:86} event structures.
We plan to use them to develop mechanised \emph{denotational} semantics 
of concurrent languages and relaxed shared memory models~\cite{Dodds-al:ESOP18}.

>>>>

First, we plan to apply our library to a wider range of problems. 
We are going to develop a mechanised semantics of some long-established
languages used to model concurrency, in particular,
the calculus of communicating systems~(CCS)~\cite{Milner:80} and 
$\pi$-calculus~\cite{Milner:99}.      
We also plan to continue our work on 
expressing various relaxed models 
of shared memory~\cite{Lahav-al:POPL16, Lahav-al:PLDI17, Podkopaev-al:POPL19}
in terms of event structures.

Second, we want to cover other classes of event structures in our library, 
in particular, bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991},
and stable~\cite{Winskel:82, Winskel:86} event structures.
We plan to use them to develop mechanised \emph{denotational} semantics 
of concurrent languages and relaxed shared memory models~\cite{Dodds-al:ESOP18}.

### 2021-05-21 papers/SYRCoSE/regmachine.tex
typo: single-threaded
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

* feat:

  r      &\in& \Reg     &   &                    & \text{thread-local register}   \\ 
  x      &\in& \Loc     &   &                    & \text{shared memory location}  \\ 
  v      &\in& \Z       &   &                    & \text{value}                   \\
  \otimes&\in& \BinOp   &   &                    & \text{binary operation}        \\
  \ip{i} &\in& \N       &   &                    & \text{instruction label}       \\

\end{array}
\] 
\caption{Syntax of the register machine}
\label{fig:regmachine-syntax}
\end{figure}

We first present the semantics of a single thread program.
Under this semantics, memory access instructions 
do not operate on shared memory
but rather produce a \emph{label}
denoting the side-effect of the operation 
(see \cref{fig:label-syntax}).
This encoding allows us to decouple 
the semantics of the register machine 
from a \emph{memory model}.

>>>>

  r      &\in& \Reg     &   &                    & \text{thread-local register} \\ 
  x      &\in& \Loc     &   &                    & \text{shared memory }        \\
         &   &          &   &                    & \qquad  \text{location}      \\
  v      &\in& \Z       &   &                    & \text{value}                 \\
  \otimes&\in& \BinOp   &   &                    & \text{binary operation}      \\
  \ip{i} &\in& \N       &   &                    & \text{instruction label}     \\

\end{array}
\]

\caption{Syntax of the register machine}
\label{fig:regmachine-syntax}
\end{figure}

We first present the semantics of a single-threaded program.
Under this semantics, memory access instructions 
do not operate on shared memory
but rather produce a \emph{label}
denoting the side-effect of the operation 
(see \cref{fig:label-syntax}).
This encoding allows us to decouple 
the semantics of the register machine 
from a \emph{memory model}.

### 2021-05-21 papers/SYRCoSE/related.tex
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

* feat: add part of small-step

More recently, event structures were applied 
in the context of relaxed memory models~%
\cite{Jeffrey-Riely:LICS16, PichonPharabod-Sewell:POPL16, Chakraborty-Vafeiadis:POPL19, Moiseenko-al:ECOOP20}.
Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
that was accompanied by a mechanisation in a proof assistant. 
The authors formalised the \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
memory model in \coq. However, this memory model uses 
a custom variant of event structures, that does not 
obey axioms of any conventional class of 
event structures~\cite{Winskel:82, Langerak:91, Boudol-Castellani:1991}. 
This fact makes it harder to reuse and adapt it 
to other applications of the theory.

>>>>

More recently, event structures were applied 
in the context of relaxed memory models~%
\cite{Jeffrey-Riely:LICS16, PichonPharabod-Sewell:POPL16, Chakraborty-Vafeiadis:POPL19, Moiseenko-al:ECOOP20}.
Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
that was accompanied by a mechanisation in a proof assistant. 
The authors formalised the \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
memory model in \coq. However, this memory model uses 
a custom variant of event structures, that does not 
obey the axioms of any conventional class of 
event structures~\cite{Winskel:82, Langerak:91, Boudol-Castellani:1991}. 
This fact makes it harder to reuse and adapt it 
to other applications of the theory.

### 2021-05-21 papers/SYRCoSE/small_step.tex
SYRCoSE: WIP --- reworking text
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

\subsection{Construction small-step semantics}
Consider we have some Execution Event Structure $es = \la E, \ffun{lab}, \fpred, \frf \ra$, $pr \in E$ ~--- the $es$'s event and some fresh identifier $e = fresh(E)$. Let's define expansion $es$ with $e$ depending on it's label:
\begin{definition}[Add event]
  \begin{itemize}
    \item $es + \la \wlab{x}{i}, \ pr \ra \ \triangleq \vspace{1mm} \\ \ 
    \la E \cup \{e\}, \
        \ffun{lab}[e \leftarrow \wlab{x}{i}], \
        \fpred[e \leftarrow pr], \
        \frf \ra\ $
        \vspace{1mm}
    \item $es + \la \text{ThreadStart} \ra \ \triangleq \vspace{1mm} \\
    \la E \cup \{e\}, \
    \ffun{lab}[e \leftarrow \text{ThreadStart}], \
    \fpred, \
    \frf \ra\ $
    \vspace{1mm}
    \item $es + \la \rlab{x}{i}, \ pr, \ wr \ra \ \triangleq$
    $$\la E \cup \{e\}, \
    \ffun{lab}[e \leftarrow \rlab{x}{i}], \
    \fpred[e \leftarrow pr], \
    \frf[e \leftarrow wr] \ra\ $$
  \end{itemize}
\input{fig/small_step.tex}
\end{definition}
Then we define programm configuration:
\begin{definition}[configuration]
  $$config \ \triangleq \ \la \es, \tmap \ra$$
  where
  \begin{itemize}
    \item $\es$ ~--- our current consistent Execution Event Structure
    \item $\tmap$ ~--- map that takes event $e$ from $\es$ and returns some pair. First component of this pair is a state of some thread after execution of $e$'s memory-side-effect. Our programm contains several threads, so we need to specify number of thread (where everything is going on) in the second component
  \end{itemize}
\end{definition}
So programm configuration describes current Execution Event Structure, where we map each event to corresponding thread state. Thus, to make our semantics step we should select some event and make progress in it's thread state. Before putting it formally we will need some notations:

By $\es.E$, $\es.\ffun{lab}$, $\es.fr$ we denote $E$, $\ffun{lab}$ and $fresh(E)$ respectively, where $\es = \la E, \ffun{lab}, \fpred, \frf \ra$. Also by $\es.tid$ we denote the smallest number of thread that wasn't started yet. \\
If we have some $p : \texttt{P}$, and $t \in \mathbb{N}$ ~--- thread number, $p[t] :=$ $t$'s thread of programm $p$.

@@

One can find definition of small-step semantics in \cref{fig:smallstep} \\
First look at $\to$ThreadStart rule. In this rule we are starting a new thread by adding some fresh identifier to $\es.E$ with label -- ThreadStart. \\
If we proceeded in thread state of some event $e$ without any memory-side-effect, we should just update $e$'s thread state (see $\to\_$). \\
Another possibility of memory-side-effect is write to some shared-memory location. In this case we should just add a new write-event to our Execution Event Structure, and map it to the thread state after our side-effect (see $\to\wlab{x}{i}$). \\
We observe the most interesting situation when our side-effect was $\rlab{x}{\_}$ (see $\to\rlab{x}{i}$). That means that we wanted to read from some shared-memory location. We need to choose some write-event $w \in \es.E$ and read value from it. It's clear that $w$ should have label $\wlab{x}{i}$, for some $i$. \\
Unfortunately, in the last step we could read from conflicting event $w$. So after making $\to$ step, it is necessary to chek if our final $\es'$ is consistent (see small-step rule). After all, we obtain one-step relation $\Rightarrow$.

>>>>

\subsection{Construction small-step semantics}
Consider we have some Execution Event Structure $es = \la E, \ffun{lab}, \fpred, \frf \ra$, $pr \in E$ ~--- the $es$'s event and some fresh identifier $e = fresh(E)$. Let's define expansion $es$ with $e$ depending on it's label:
\begin{definition}[Add event]
  \begin{itemize}
    \item $es + \la \wlab{x}{i}, \ pr \ra \ \triangleq \vspace{1mm} \\ \ 
    \la E \cup \{e\}, \
        \ffun{lab}[e \leftarrow \wlab{x}{i}], \
        \fpred[e \leftarrow pr], \
        \frf \ra\ $
        \vspace{1mm}
    \item $es + \la \text{ThreadStart} \ra \ \triangleq \vspace{1mm} \\
    \la E \cup \{e\}, \
    \ffun{lab}[e \leftarrow \text{ThreadStart}], \
    \fpred, \
    \frf \ra\ $
    \vspace{1mm}
    \item $es + \la \rlab{x}{i}, \ pr, \ wr \ra \ \triangleq$
    $$\la E \cup \{e\}, \
    \ffun{lab}[e \leftarrow \rlab{x}{i}], \
    \fpred[e \leftarrow pr], \
    \frf[e \leftarrow wr] \ra\ $$
  \end{itemize}

\end{definition}
Then we define program configuration:
\begin{definition}[configuration]
  $$config \ \triangleq \ \la \es, \tmap \ra$$
  where
  \begin{itemize}
    \item $\es$ ~--- our current consistent Execution Event Structure
    \item $\tmap$ ~--- map that takes event $e$ from $\es$ and returns some pair. First component of this pair is a state of some thread after execution of $e$'s memory-side-effect. Our program contains several threads, so we need to specify number of thread (where everything is going on) in the second component
  \end{itemize}
\end{definition}
So program configuration describes current Execution Event Structure, where we map each event to corresponding thread state. Thus, to make our semantics step we should select some event and make progress in it's thread state. Before putting it formally we will need some notations:

By $\es.E$, $\es.\ffun{lab}$, $\es.fr$ we denote $E$, $\ffun{lab}$ and $fresh(E)$ respectively, where $\es = \la E, \ffun{lab}, \fpred, \frf \ra$. Also by $\es.tid$ we denote the smallest number of thread that wasn't started yet. \\
If we have some $p : \texttt{P}$, and $t \in \mathbb{N}$ ~--- thread number, $p[t] :=$ $t$'s thread of program $p$.

@@

One can find definition of small-step semantics in \cref{fig:smallstep} \\
First look at $\to$ThreadStart rule. In this rule we are starting a new thread by adding some fresh identifier to $\es.E$ with label -- ThreadStart. \\
If we proceeded in thread state of some event $e$ without any memory-side-effect, we should just update $e$'s thread state (see $\to\_$). \\
Another possibility of memory-side-effect is write to some shared-memory location. In this case we should just add a new write-event to our Execution Event Structure, and map it to the thread state after our side-effect (see $\to\wlab{x}{i}$). \\
We observe the most interesting situation when our side-effect was $\rlab{x}{\_}$ (see $\to\rlab{x}{i}$). That means that we wanted to read from some shared-memory location. We need to choose some write-event $w \in \es.E$ and read value from it. It's clear that $w$ should have label $\wlab{x}{i}$, for some $i$. \\
Unfortunately, in the last step we could read from conflicting event $w$. So after making $\to$ step, it is necessary to check if our final $\es'$ is consistent (see small-step rule). After all, we obtain one-step relation $\Rightarrow$.
